---
title: 'Tooth Growth Analysis: Correlation with diet and Vitamin C'
author: "Winston"
date: "October 21, 2014"
output: word_document
---

####Note to reader
This .pdf may look at bit strange because just as I was finishing I upgraded my Mac OS to Yosemite. The LaTeX version seemed to stop working with RPubs. So I saved as a word file and the saved that as a .pdf, which I uploaded.

##Get the data and take a quick look at it

```{r}
data(ToothGrowth)
```



```{r, echo=FALSE}
str(ToothGrowth)
```

##Exploratory Analysis
Here is a plot of the data showing behavior trends in tooth growth factored by supplement.

```{r "plot", warning=FALSE, fig.height=3, fig.width=4}
require(ggplot2)

ggplot(ToothGrowth, aes(x=dose, y = len))+geom_point(size = 5, colour="red")+geom_point(size = 4, colour="lightblue")+facet_grid(. ~ supp)


```

A summary of the data:  
```{r "summary", echo=FALSE}
summary(ToothGrowth)
```  
shows the data appear to be well behaved (10 unpaired observations at each condition) and do not require cleaning for this anlaysis.

```{r "dataheaderX", echo=FALSE}
###display the header of the data 
head(ToothGrowth, n=5)
```

##Confidence Interval Analysis
The first step is to look at the statistical significance of the shift in the data with increasing dose. The conclusion after looking at all the pairs is that there is an upward trend (meaning higher doses correlate to statistically significant higher length) among all the data.  

To keep the analysis part clean, first split up the data by Supplement and Dose

```{r "split up data"}

##Group the data into individual sets with somewhat descriptive names (Supplement and Dose)
        ##Vitamin C set
        VC05<-ToothGrowth[1:10,]
        VC10<-ToothGrowth[11:20,]
        VC20<-ToothGrowth[21:30,]
        ##Ornage Juice Set
        OJ05<-ToothGrowth[31:40,]
        OJ10<-ToothGrowth[41:50,]
        OJ20<-ToothGrowth[51:60,]

```

##Within Vitamin C Data
Some smaple anlaysis

####Dose0.5 to Dose1.0

```{r "compare1"}

a<- t.test(VC10$len, VC05$len, paried=FALSE)
lcb<-a$conf.int[1]
ucb<-a$conf.int[2]

```

The confidence interval differents between tooth length at these doses is `r lcb` to `r ucb` which `r if(ucb-lcb>abs(ucb)-abs(lcb)) {"does"} else {"does not"}` contain 0.  
   
####Dose1.0 to Dose2.0

```{r "compare2", echo=FALSE}

a<- t.test(VC20$len, VC10$len, paried=FALSE)
lcb<-a$conf.int[1]
ucb<-a$conf.int[2]

```

The confidence interval difference between tooth length at these doses is `r lcb` to `r ucb` which `r if(ucb-lcb>abs(ucb)-abs(lcb)) {"does"} else {"does not"}` contain 0.


##Within Orange Juice
We can draw a similar conclusion that the correlation of increased does and tooth length is significnat statistically.  

####Dose0.5 to Dose1.0

```{r "compare4"}

a<- t.test(OJ10$len, OJ05$len, paried=FALSE)
lcb<-a$conf.int[1]
ucb<-a$conf.int[2]

```

The confidence interval difference between tooth length at these doses is `r lcb` to `r ucb` which `r if(ucb-lcb>abs(ucb)-abs(lcb)) {"does"} else {"does not"}` contain 0.

####Dose1.0 to Dose1.0

```{r "compare7"}

a<- t.test(OJ10$len, VC10$len, paried=FALSE)
lcb<-a$conf.int[1]
ucb<-a$conf.int[2]

```

The confidence interval difference between tooth length from the Orange Juice and Vitamin C at doses of 1.0 is `r lcb` to `r ucb` which `r if(ucb-lcb>abs(ucb)-abs(lcb)) {"does"} else {"does not"}` contain 0.


####Conclusion  
In all the cases above the shifts in the mean, both with increasing dose and the comparison of Vitamin C to Orange Juice are significant. 


##Hypothesis Testing

###Hypothesis to be tested. 

Orange Juice is more effective (produces more tooth growth) than Vitamin C.

####Test at dose of 1.0.  

```{r "hypothesis"}
b<-t.test(OJ10$len, VC10$len, paired=FALSE, var.equal=TRUE)

```

The anlaysis shows that for `r substring(b$data.name,1,4)` and `r substring(b$data.name,14,17)` using the `r b$method` the t-statistic is `r b$parameter` and the confidence interval is `r b$conf.int[1]` and `r b$conf.int[2]`.

####Test at dose of 2.0

```{r "hypothesis2"}
b<-t.test(OJ20$len, VC20$len, paired=FALSE, var.equal=TRUE)

```

The anlaysis shows that for `r substring(b$data.name,1,4)` and `r substring(b$data.name,14,17)` using the `r b$method` the t-statistic is `r b$parameter` and the confidence interval is `r b$conf.int[1]` and `r b$conf.int[2]`.

##Assumptions

Here I assume that the variances of the distributions are equal between the Orange Juice and the Vitamin C populations. 

##Conclusion
The hypothesis that Orange Juice is more effective than Vitamin C in tooth length increases is **FALSE** since at higher doses the difference in the outcome includes zero. 

However, at lower doses, it does appear Orange Juice is more effective. 



