---
title: "Tooth Growth Analysis: Correlation with diet and Vitamin C"
author: "Winston"
date: "October 21, 2014"
output: pdf_document
---

##Get the data and take a quick look at it

```{r}
data(ToothGrowth)
```



```{r, echo=FALSE}
str(ToothGrowth)
```

##Exploratory Analysis
Here is a plot of the data showing behavior trends in tooth growth factored by supplement.

```{r "plot"}
require(ggplot2)

ggplot(ToothGrowth, aes(x=dose, y = len))+geom_point(size = 2)+facet_grid(. ~ supp)+geom_smooth()


```

A summary of the data:  
`r summary(ToothGrowth)`  
shows the data appear to be well behaved (10 unpaired observations at each condition) and do not require cleaning for this anlaysis.

`r head(ToothGrowth)

##Confidence Intervals
The first step is to look at the statistical significance of the shift in the data. There are six pair-wise comparisons that make sense, three for each supplement.

To keep the analysis part clean, first split up the data by Supplement and Dose

```{r "split up data"}

##Group the data into individual sets with somewhat descriptive names (Supplement and Dose)
        ##Vitamin C set
        VC05<-ToothGrowth[1:10,]
        VC10<-ToothGrowth[11:20,]
        VC20<-ToothGrowth[21:30,]
        ##Ornage Juice Set
        OJ05<-ToothGrowth[31:40,]
        OJ10<-ToothGrowth[41:50,]
        OJ20<-ToothGrowth[51:60,]

```

###Vitamin C

####Dose0.5 to Dose1.0

```{r "compare1"}

a<- t.test(VC10$len, VC05$len, paried=FALSE)
lcb<-a$conf.int[1]
ucb<-a$conf.int[2]

```

The confidence interval differents between tooth length at these doses is `r lcb` to `r ucb` which `r if(ucb-lcb>abs(ucb)-abs(lcb)) {"does"} else {"does not"}` contain 0.

####Dose1.0 to Dose2.0

```{r "compare2"}

a<- t.test(VC20$len, VC10$len, paried=FALSE)
lcb<-a$conf.int[1]
ucb<-a$conf.int[2]

```

The confidence interval differents between tooth length at these doses is `r lcb` to `r ucb` which `r if(ucb-lcb>abs(ucb)-abs(lcb)) {"does"} else {"does not"}` contain 0.

####Dose0.5 to Dose2.0

```{r "compare3"}

a<- t.test(VC20$len, VC05$len, paried=FALSE)
lcb<-a$conf.int[1]
ucb<-a$conf.int[2]

```

The confidence interval differents between tooth length at these doses is `r lcb` to `r ucb` which `r if(ucb-lcb>abs(ucb)-abs(lcb)) {"does"} else {"does not"}` contain 0.


###Orange Juice

####Dose0.5 to Dose1.0

```{r "compare4"}

a<- t.test(OJ10$len, OJ05$len, paried=FALSE)
lcb<-a$conf.int[1]
ucb<-a$conf.int[2]

```

The confidence interval differents between tooth length at these doses is `r lcb` to `r ucb` which `r if(ucb-lcb>abs(ucb)-abs(lcb)) {"does"} else {"does not"}` contain 0.

####Dose1.0 to Dose2.0

```{r "compare5"}

a<- t.test(OJ20$len, OJ10$len, paried=FALSE)
lcb<-a$conf.int[1]
ucb<-a$conf.int[2]

```

The confidence interval differents between tooth length at these doses is `r lcb` to `r ucb` which `r if(ucb-lcb>abs(ucb)-abs(lcb)) {"does"} else {"does not"}` contain 0.

####Dose0.5 to Dose2.0

```{r "compare6"}

a<- t.test(OJ20$len, OJ05$len, paried=FALSE)
lcb<-a$conf.int[1]
ucb<-a$conf.int[2]

```

The confidence interval differents between tooth length at these doses is `r lcb` to `r ucb` which `r if(ucb-lcb>abs(ucb)-abs(lcb)) {"does"} else {"does not"}` contain 0.